---
title: "Health Insurance and Data Science"
author: "Miguel Briones"
date: "3/27/2018"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(shiny)
library(rpart)
library(tidyverse)
library(rpart.plot)
library(caret)
library(dplyr)
library(scales)
library(ggplot2)
```

## Challenge

- Insurance company needed to predict the cost of insurance payouts based on injury type and location of accident.


## Action

- Take data from previous insurance payouts and try to predict average cost 


## Results - Frequency of each Injury type

```{r cars, echo = FALSE}
data <- read.csv("Data/Insurance/Table.csv", stringsAsFactors = FALSE)
names(data)[names(data) == "DAÑO"] <- "DANO"

data <- data.frame(lapply(data, function(x) {
          gsub("ñ", "n", x)
              }))

data <- data.frame(lapply(data, function(x) {
          gsub("ó", "o", x)
              }))

data <- data.frame(lapply(data, function(x) {
          gsub("á", "a", x)
              }))

data <- data.frame(lapply(data, function(x) {
          gsub("é", "e", x)
              }))

data <- data.frame(lapply(data, function(x) {
          gsub("Í", "I", x)
              }))

data <- data.frame(lapply(data, function(x) {
          gsub("Ñ", "N", x)
              }))

data <- data.frame(lapply(data, as.character), stringsAsFactors=FALSE)

dat <- data %>%
  select(-1) %>%
  filter(!is.na(IMPORTE)) %>% #Remove rows where IMPORTE is NA
  filter(COMUNIDAD_AUTONOMA != "") %>%
  select(c(DANO,ESPECIALIDAD,LUGAR,CENTRO,IMPORTE,COMUNIDAD_AUTONOMA))
dat <- dat[!(dat$DANO == "No Definido" | dat$DANO == "Desconocido"), ]
dat$IMPORTE <- as.numeric(dat$IMPORTE)
dat <- dat[dat$IMPORTE >= 0, ]

dat_dano <- dat %>%
  group_by(DANO) %>%
  summarise(count = n(), cost_mean = mean(IMPORTE)) %>%
  arrange(desc(cost_mean))

ggplot(dat_dano) +
  geom_col(aes(x=DANO, y=count)) +
  coord_flip() +
  ggtitle("Frequency of each DAÑO value")
```


## Results - Mean Cost for each Injury type
```{r, echo=FALSE}
ggplot(dat_dano) +
  geom_col(aes(x=DANO, y=cost_mean)) +
  coord_flip() +
  ggtitle("Mean IMPORTE for each DAÑO value")
```

## Interpretation of Injury Type

From these plots, it looks like Osteomuscular, Muerte, and Daño Material are the most frequent types of injury, and Neurológico grave, Daño al niño o al recién nacid, Coma are the most expensive types of injury. It is understandable that these severe injuries are more expensive and less frequent.

## Results - Frequency of each Location
```{r, echo=FALSE}
dat_ca <- dat %>%
  group_by(COMUNIDAD_AUTONOMA) %>%
  summarise(count = n(), cost_mean = mean(IMPORTE)) %>%
  arrange(desc(cost_mean))

ggplot(dat_ca) +
  geom_col(aes(x=COMUNIDAD_AUTONOMA, y=count)) +
  coord_flip() +
  ggtitle("Frequency of each COMUNIDAD_AUTONOMA value")
```

## Results - Mean Cost at Each Location
```{r, echo = FALSE}
ggplot(dat_ca) +
  geom_col(aes(x=COMUNIDAD_AUTONOMA, y=cost_mean)) +
  coord_flip() +
  ggtitle("Mean IMPORTE for each COMUNIDAD_AUTONOMA value")
```

## Interpretation
The average claim cost varies from region to region. 

## Basic Machine Learning
We used Classification Trees, which is a machine learning method used to predict categorical outcomes, to model and predict the cost level (high, low) based on the injury type (DAÑO) and region (COMUNIDAD_AUTONOMA).
```{r, echo = FALSE}
dat$log <- log10(dat$IMPORTE)

dat_min <- dat %>%
  filter(dat$IMPORTE <= 3) %>%
  mutate(cost = 'Bajo')
dat_hi <- dat %>%
  filter(dat$IMPORTE >3)%>%
  mutate(cost = 'Alto')
datex <- bind_rows(dat_min, dat_hi)

set.seed(101)
alpha     <- 0.7 # percentage of training set
datfac <- datex
datfac$DANO <- factor(datfac$DAN)
datfac$COMUNIDAD_AUTONOMA <- factor(datfac$COMUNIDAD_AUTONOMA)
datfac$LUGAR <- factor(datfac$LUGAR)
datfac$ESPECIALIDAD <- factor(datfac$ESPECIALIDAD)

datfac$cost <- factor(datfac$cost)

inTrain   <- sample(1:nrow(datfac), alpha * nrow(datfac))
train.set <- datfac[inTrain,]
test.set  <- datfac[-inTrain,]

fit <- rpart(cost ~ DANO + COMUNIDAD_AUTONOMA,
             method="class", data=train.set) #it will take a few min

rpart.plot(fit)
```

## Testing our model
```{r, echo = TRUE}
maxidx <- function(arr) {
  return(which(arr == max(arr)))
}


pred <- predict(fit,test.set)
idx <- apply(pred, c(1), maxidx)
prediction <- c('high', 'low')[idx]
confMat <- table(test.set$cost, prediction)
accuracy <- sum(diag(confMat))/sum(confMat)
accuracy
```
Accuracy rate = 85.4%. These results can be dramaticly improved with better data and more time. 

## Shiny App

```{r, echo=FALSE}
datains <- read.csv("Data/Insurance/datalog.csv")

fit <- rpart(cost ~ DANO + COMUNIDAD_AUTONOMA,
               method="class", data=datains)

shinyApp(
  ui <- shinyUI(fluidPage(
    
    br(),
    
    h1("Insurance Claim Prediction", align = "center"),
    br(),
    h5("Please Select the Injury Type, Followed by The Location", align = "center"),
    
    br(),
    br(),
    
    fluidRow(align = "center", 
             column(5, selectInput('Injury', 'Injury Type', choices = as.factor(unique(datains$DANO)), selected =    "Digestivo y boca")
             ),
             column(7, selectInput('Location', 'Location', ""))
    ),
    
    hr(),
    
    fluidRow(align = "center", 
             column(5, strong("Predicted Cost:"), textOutput("costs")
             ),
             column(7, strong("Probability:"), textOutput("Prob"))
    ),
    hr(),
    
    fluidRow(align = "center", 
             column(5, strong("AverageCost:"), textOutput("Average")
             ),
             column(7, strong("+/-:"), textOutput("Uncertain"))
    ),
    hr(),
    
    fluidRow(align = "center", 
             column(12, strong("Outlier:"), textOutput("Outlier"))
    ),
    hr(),
    
    h5("Where's the Average Cost?", align = "center"),
    br(),
    p("The dotted line represents the average cost within the known range of costs 
       based on historcal data.The shaded area represents the uncertainty that 
      surrounds the average cost based on injury type and location"),
    
    plotOutput("histPlot"),
    
    hr()
    )
  ),
  
  # Define server logic 
  server <- shinyServer(function(input, output,session) {
    
    outVar <- reactive({
      newdatains <- datains %>% filter(DANO == input$Injury & IMPORTE != 0)
      unique(newdatains$COMUNIDAD_AUTONOMA)
    })
    observe({
      updateSelectInput(session, "Location", choices = outVar())
    })
    
    modelpred1 <- reactive({
      Injurycosts <- input$Injury
      Locationcosts <- input$Location
      as.character(predict(fit, newdata = data.frame(DANO = Injurycosts, COMUNIDAD_AUTONOMA = Locationcosts), 
                           type = "class"))
    })
    
    modelpred2 <- reactive({
      Injurycosts <- input$Injury
      Locationcosts <- input$Location
      percent(max(predict(fit, newdata = data.frame(DANO = Injurycosts, COMUNIDAD_AUTONOMA = Locationcosts))))
    })
    
    importecos <- reactive({
      sapply(subset(datains, DANO == input$Injury & COMUNIDAD_AUTONOMA == input$Location, 
                    select = c(IMPORTE)), mean)
    })
    
    importeunc <- reactive({
      importelen <- subset(datains, DANO == input$Injury & COMUNIDAD_AUTONOMA == input$Location, 
                           select = c(IMPORTE))
      sd <- sapply(subset(datains, DANO == input$Injury & COMUNIDAD_AUTONOMA == input$Location, 
                          select = c(IMPORTE)), sd)
      elen <- sd/sqrt(nrow(importelen))
      ifelse(is.na(elen) == TRUE, "No Hay Suficientes Puntos de Datos", elen)
    })
    
    importeout <- reactive({
      importelen2 <- subset(datains, DANO == input$Injury & COMUNIDAD_AUTONOMA == input$Location, 
                            select = c(IMPORTE))
      importemean <- sapply(subset(datains, DANO == input$Injury & COMUNIDAD_AUTONOMA == input$Location, 
                                   select = c(IMPORTE)), mean)
      importebox <- boxplot(importelen2, plot=FALSE)$out #outside 1.5 times the interquartile range above the upper quartile and below the lower quartile
      ifelse(importemean %in% importebox, "Yes", "No")
    })
    
    output$costs <- renderText({(modelpred1())
    })
    
    output$Prob <- renderText({(modelpred2())
    })
    
    output$Average <- renderText({importecos() 
    })
    
    output$Uncertain <- renderText({importeunc() 
    })
    
    output$Outlier <- renderText({as.character(importeout()) 
    })
    
    output$histPlot <- renderPlot({
      plotcost <- as.numeric(importecos())
      plotunc <- as.numeric(importeunc())
      
      importeplus <- plotcost + plotunc
      importeminus <- plotcost - plotunc
      
      density_plot <- ggplot(datains, aes(log)) +
        geom_density(binwidth = 0.5) +
        #xlim(0,6) + 
        scale_x_continuous(breaks = 0:6,
                           labels = c("$1", "$10", "$100", "$1,000", "$10,000", "$100,000", "$1,000,000")) +
        xlab("") +
        ylab("") +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
              panel.background = element_blank(), axis.line = element_blank(), axis.text.x = element_text(),
              axis.text.y=element_blank(),
              axis.ticks.y=element_blank()) 
      
      dpb <- ggplot_build(density_plot)
      
      x1 <- min(which(dpb$data[[1]]$x >= log10(importeplus)))
      x2 <- max(which(dpb$data[[1]]$x <= log10(importeminus)))
      
      density_plot + geom_area(data=data.frame(x=dpb$data[[1]]$x[x1:x2],
                                               y=dpb$data[[1]]$y[x1:x2]),
                               aes(x=x, y=y), fill = "light blue") + geom_vline(aes(xintercept=log10(plotcost)),
                                                                                color="black", linetype="dashed",   size=0.5)
    })
  })
)

```

## Further Applications

- Personalized Medicine
- Genomics
- Predictive Analytics and Preventive Measures
- Patient Monitoring and Home Devices
- Disease Modeling and Mapping


