---
title: "People Analytics"
author: "Miguel Briones"
date: "3/27/2018"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(flexdashboard)
library(ggrepel)
library(shiny)
library(ggplot2)
library(plotly)
library(plyr)
library(tidyverse)
library(leaflet)
library(maps)
library(scales)
library(sp)
library(rgdal)
library(maptools)
library(httr)
library(broom)
library(ggmap)
library(wordcloud)
```

## What is People Analytics

- A data-driven approach to inform your people practices, programs and processes.

- Analytical techniques, ranging from reporting and metrics to predictive analytics to experimental research can help you uncover new insights, solve people problems and direct your HR actions.

- A foundational building block that informs actions and ultimately keep employers 


## Examples of People Analytics
- Knowing the length of time it takes to fill a position can help you assess the efficiency of your recruiting process.

- Knowing your turnover rate can be a useful indicator of your organizational culture, health, and finances.

- A career path  metric can help break out your various career ladders by measuring how many internal moves are   promotions versus lateral transfers across different departments and specialties.

- Knowing the revienue per employee statistics can help you determine how over or understaffed your organization might be by examining the growth of your revenue over time.

## Example - Where do Queens College students end up?

Queens College needed to monitor the value of their degrees. The value of a STEM degree continues to rise, while the value of a humanities degree declines. Queens College wanted to understand how two specific degrees, Computer Science & English, fared on the job market.

## Methods of Action

- Scraped LinkedIn Profiles for Queens College graduates to gain insight into job and job history

- Scraped Glassdoor for salary related to current job for Queens College graduate

## Results - Computer Science in the Workforce
```{r, echo = FALSE}
cuny <- read.csv("Data/Queens/cunylinkedin.csv", stringsAsFactors = F)
glassdravg <- read.csv("Data/Queens/glassdooraveragesalary.csv", stringsAsFactors = F)
glassdrsal <- read.csv("Data/Queens/glassdoorsalary.csv", stringsAsFactors = F)

cuny$AvgSalary <- glassdravg$AverageSalary[match(cuny$Job, glassdravg$Job)]
cuny$Salary <- glassdrsal$Salary[match(cuny$JobCompany, glassdrsal$JobCompany)]
cuny$Sector[is.na(cuny$Salary)] <- "Unemployed/Not Found"

cuny$Salary[grepl("PHD", cuny$JobTitle)] <- 35530
cuny$Salary[grepl("PhD", cuny$JobTitle)] <- 35530
cuny$Salary[grepl("Doctoral", cuny$JobTitle)] <- 35530
cuny$Salary[grepl("Teaching Assistant", cuny$JobTitle)] <- 35530
cuny$Salary[grepl("Graduate Student", cuny$JobTitle)] <- 35530
cuny$Salary[grepl("Unknown Company", cuny$Company)] <- NA

cuny$Salary[grepl("NYC Department of Education", cuny$Company)] <- 58600
cuny$Sector[grepl("NYC Department of Education", cuny$Company)] <- "Public Education"

topskillsQueensCS <- read.csv("Data/Queens/topskillsQueensCS.csv")
topskillsQueensCS[c("X")] <- list(NULL)
topskillsQueensCS <- subset(topskillsQueensCS, select=c(Skills, Count))
colnames(topskillsQueensCS) <- c("Skills", "Number of Graduates")

topskillsQueensEN <- read.csv("Data/Queens/topskillsQueensEN.csv")
topskillsQueensEN[c("X")] <- list(NULL)
topskillsQueensEN <- subset(topskillsQueensEN, select=c(Skills, Count))
colnames(topskillsQueensEN) <- c("Skills", "Number of Graduates")

sectorcount <- subset(cuny, Degree == "Computer Science" & School == "Queens College", select = c(Sector, Salary))
sectorcount$Count <- rep(1)
sectorcount <- aggregate(Count ~ Sector, data = sectorcount, sum)
salarysector <- subset(cuny, Degree == "Computer Science" & School == "Queens College", select = c(Sector, Salary))
salarysector <- aggregate(Salary ~ Sector, data = salarysector, mean)
sectorcount$Salary <- salarysector$Salary[match(sectorcount$Sector, salarysector$Sector)]

QueensPieCS <- plot_ly(sectorcount, labels = ~Sector, values = ~Count, type = 'pie',textposition = 'outside',textinfo = 'label+percent', hoverinfo = 'text', text = ~paste('Average Salary of $', round(Salary))) %>%
  layout(title = 'Queens College Computer Science',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

QueensPieCS
```

## Results - English in the Workforce
```{r, echo=FALSE}
sectorENcount <- subset(cuny, Degree == "English" & School == "Queens College", select = c(Sector, Salary))
sectorENcount$Count <- rep(1)
sectorENcount <- aggregate(Count ~ Sector, data = sectorENcount, sum)
salaryENsector <- subset(cuny, Degree == "English" & School == "Queens College", select = c(Sector, Salary))
salaryENsector <- aggregate(Salary ~ Sector, data = salaryENsector, mean)
sectorENcount$Salary <- salaryENsector$Salary[match(sectorENcount$Sector, salaryENsector$Sector)]

QueensPieEN <- plot_ly(sectorENcount, labels = ~Sector, values = ~Count, type = 'pie',textposition = 'outside',textinfo = 'label+percent', hoverinfo = 'text', text = ~paste('Average Salary of $', round(Salary))) %>%
  layout(title = 'Queens College English',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

QueensPieEN
```


## Results - Value of Queens College Degree
```{r, echo=FALSE}
Queenssalaryyear <- subset(cuny, School == "Queens College" & (!is.na(Salary)), 
                              select = c(Degree, AvgSalary, YearGrad, Salary))

     
Queenssalaryyear2 <- Queenssalaryyear %>% group_by(Degree, YearGrad) %>% summarize(Current = mean(Salary))

Queenssalaryyear3 <- Queenssalaryyear %>% group_by(Degree, YearGrad) %>% summarize(Average = mean(AvgSalary))

ComputerSciCurrent <- subset(Queenssalaryyear2, Degree == "Computer Science", select = c(YearGrad, Current))

ComputerSciAverage <- subset(Queenssalaryyear3, Degree == "Computer Science", select = c(YearGrad, Average))

EnglishCurrent <- subset(Queenssalaryyear2, Degree == "English", select = c(YearGrad, Current))

EnglishAverage <- subset(Queenssalaryyear3, Degree == "English", select = c(YearGrad, Average))

colnames(ComputerSciCurrent) <- c("YearGrad", "CurrentCS")

colnames(ComputerSciAverage) <- c("YearGrad", "AverageCS")

colnames(EnglishCurrent) <- c("YearGrad", "CurrentEN")

colnames(EnglishAverage) <- c("YearGrad", "AverageEN")

dataplotly <- Reduce(function(x, y) merge(x, y, all=TRUE), list(ComputerSciAverage, ComputerSciCurrent, EnglishAverage, EnglishCurrent))



Queenssalaryyearplot <- plot_ly(dataplotly, x = ~YearGrad, y = ~CurrentCS, name = 'CS Current Salary', type = 'scatter', mode = 'lines', line = list(color = 'rgb(0, 0, 128)', width = 4)) %>%
  add_trace(y = ~AverageCS, name = 'CS Average Salary', line = list(color = 'rgb(65,105,225)', width = 4, dash = 'dash')) %>%
  add_trace(y = ~CurrentEN, name = 'EN Current Salary', line = list(color = 'rgb(139,0,0)', width = 4)) %>%
  add_trace(y = ~AverageEN, name = 'EN Average Salary', line = list(color = 'rgb(250,128,114)', width = 4, dash = 'dash')) %>%
  layout(title = "Salaries for Queens College Graduates",
         xaxis = list(title = "Year Graduated"),
         yaxis = list (title = "Salary (In U.S. Dollars)"))


Queenssalaryyearplot 
```

## Results - Top Skills for Computer Science Graduates
```{r, echo=FALSE}
topskillsQueensCS$`Number of Graduates` <- gsub(",", "", topskillsQueensCS$`Number of Graduates`)

topskillsQueensCS$`Number of Graduates`<-as.character(topskillsQueensCS$`Number of Graduates`)

topskillsQueensCS$`Number of Graduates`<-as.numeric(topskillsQueensCS$`Number of Graduates`)

piequeensCSskills <- plot_ly(topskillsQueensCS, labels = ~Skills, values = ~`Number of Graduates`, type = 'pie',textposition = 'outside',textinfo = 'label') %>%
  layout(title = 'Queens CS Skills',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

piequeensCSskills
```

## Results - Top Skills for English Graduates
```{r, echo=FALSE}
topskillsQueensEN$`Number of Graduates` <- gsub(",", "", topskillsQueensEN$`Number of Graduates`)

topskillsQueensEN$`Number of Graduates`<-as.character(topskillsQueensEN$`Number of Graduates`)

topskillsQueensEN$`Number of Graduates`<-as.numeric(topskillsQueensEN$`Number of Graduates`)

piequeensENskills <- plot_ly(topskillsQueensEN, labels = ~Skills, values = ~`Number of Graduates`, type = 'pie',textposition = 'outside',textinfo = 'label') %>%
  layout(title = 'Queens EN Skills',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

piequeensENskills
```


## Further Applications
- Education: Nudging students in the right direction
- Understanding how employees can value from their benefits
- Nudging employees for health and wellness
- Nudging employees for retirement
- Hiring to fit your team  Sabermetrics for the workplace
- Fair pay practices

